---
title: "ShinyNotes"
author: "Dave"
date: "6/27/2021"
output: html_document
---

### From https://shiny.rstudio.com/tutorial/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)

ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num", 
                           label = "Choos a number", 
                           value = 25, min = 1, max = 100),
		plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output) {}

shinyApp(ui = ui, server = server)
```

All input functions use an inputId = , then label = , 

output functions:
dataTableOutput() - interactive table
htmlOutput() - raw HTML
impageOutput() - image
plotOutput() - plot
tableOutput()
testOutput()
uiOutput() - shiny UI element
verbatimTextOutput()

sintax:
plotOutput(outputId = 'hist')

server function:
this builds the outputs we want to see
There are 3 rules for the server function:
1) save objects to display to output$ Ex: output$hist <- (stuff)
2) what is saved to the output$ should use the render*() 
renderDataTable()
renderImage, Plot, Print, Table, Text, UI

Ex: 
```{r}
ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num"
                           label = "Choos a number", 
                           value = 25, min = 1, max = 100),
		          plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output){
  output$hist = renderPlot({
    title = "100 random normal values"
    hist(rnorm(100), main = title)
  })
}

shinyApp(ui = ui, server = server)
```

3) input values with input$

```{r}
ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num", # this num is accessable as input$num, changes with slinder selection
                           label = "Choos a number", 
                           value = 25, min = 1, max = 100),
		          plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output){
  output$hist = renderPlot({
    title = paste(input$num, "random normal values")
    hist(rnorm(input$num), main = title)
  })
}

shinyApp(ui = ui, server = server)
```


## share you app:
First save you app in a standard R format. 
One directory with every file the app needs (app.R, data, images, css, helper scripts, etc.)
The script must be saved as app.R. 

Other format is a two file structure. 
1) ui.R - all user interface script, everything save to the ui variable above
2) server.R - everything saved to the server variable above
Then you don't need to call shinyApp()

useful for very large apps. 

RStudio will recognize these three file names and will have a Run App button with different viewing options. 

## Shinyapps.io
free server provided by RStudio


shiny.rstudio.com/articles/shinyapps.html

Or build your own server:
www.rstudio.com/products/shiny/shiny-server/
a back end program that builds a linux web server specifically designed to host Shiny apps. 

More info and examples at shiny.rstudio.com


## End of Part 1 42:00 min



