---
title: "ShinyNotes"
author: "Dave"
date: "6/27/2021"
output: html_document
---

### From https://shiny.rstudio.com/tutorial/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shiny)

ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num", 
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
		plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output) {}

shinyApp(ui = ui, server = server)
```

All input functions use an inputId = , then label = , 

output functions:
dataTableOutput() - interactive table
htmlOutput() - raw HTML
impageOutput() - image
plotOutput() - plot
tableOutput()
testOutput()
uiOutput() - shiny UI element
verbatimTextOutput()

sintax:
plotOutput(outputId = 'hist')

server function:
this builds the outputs we want to see
There are 3 rules for the server function:
1) save objects to display to output$ Ex: output$hist <- (stuff)
2) what is saved to the output$ should use the render*() 
renderDataTable()
renderImage, Plot, Print, Table, Text, UI

Ex: 
```{r}
ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num",
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
		          plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output){
  output$hist = renderPlot({
    title = "100 random normal values"
    hist(rnorm(100), main = title)
  })
}

shinyApp(ui = ui, server = server)
```

3) input values with input$

```{r}
ui = fluidPage(
               #Input (from user)
               #example a slider function
               sliderInput(inputId = "num", # this num is accessable as input$num, changes with slinder selection
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
		          plotOutput(outputId = 'hist')
               #output (result of function returned to user)
               )

server = function(input, output){
  output$hist = renderPlot({
    title = paste(input$num, "random normal values")
    hist(rnorm(input$num), main = title)
  })
}

shinyApp(ui = ui, server = server)
```


## share you app:
First save you app in a standard R format. 
One directory with every file the app needs (app.R, data, images, css, helper scripts, etc.)
The script must be saved as app.R. 

Other format is a two file structure. 
1) ui.R - all user interface script, everything save to the ui variable above
2) server.R - everything saved to the server variable above
Then you don't need to call shinyApp()
3) the input object will update the output as the input changes

useful for very large apps. 

RStudio will recognize these three file names and will have a Run App button with different viewing options. 

## Shinyapps.io
free server provided by RStudio


shiny.rstudio.com/articles/shinyapps.html

Or build your own server:
www.rstudio.com/products/shiny/shiny-server/
a back end program that builds a linux web server specifically designed to host Shiny apps. 

More info and examples at shiny.rstudio.com


## End of Part 1 42:00 min

## Part 2
Shiny showcase
www.rstudio.com/products/shiny/shiny-user-showcase/

### Reactivity

Reactive values - ex. input$num 
They do no work on their own, they work with the reactive functions - eg. renderPlot

Reactive toolkit:
7 tools - 
render* functions: 

```{r}
ui = fluidPage(
               sliderInput(inputId = "num", 
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
               textInput(inputId = 'title', 
                         label = "random normal values", 
                         value = "Histogram of Random Normal Values"),
		          plotOutput('hist')
               )

server = function(input, output){
  output$hist = renderPlot({
    hist(rnorm(input$num), main = input$title)
  })
}

shinyApp(ui = ui, server = server)
```

### Modularize Code with reactive()
this updates reactive objects, use if you want to manipulate data and the call it by other functions downstream. 
Recative functions are functions (e.g. data() )

So reactive function notifies downstream uses are invalid. Then the output function calls the data() function. Now second fuction (stats) is invalid and run the data() function. The data() function is still valid, so only returns the value and does not rerun the data() function. 

```{r}
ui = fluidPage(
               sliderInput(inputId = "num", 
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
               textInput(inputId = 'title', 
                         label = "random normal values", 
                         value = "Histogram of Random Normal Values"),
		          plotOutput('hist'),
		          verbatimTextOutput('stats')
               )

server = function(input, output){
  data = reactive({
    rnorm(input$num)
  })
  output$hist = renderPlot({
    hist(data(), main = input$title)
  })
  output$stats = renderPrint({
    summary(data())
  })
}

shinyApp(ui = ui, server = server)
```

## Prevent reactions with isolate()
here the name field edit rename. We use isolate to prevent the refreshing of the main title when typing. Now it only refreshes the title when a new number is selected. 

```{r}
ui = fluidPage(
               sliderInput(inputId = "num", 
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
               textInput(inputId = 'title', 
                         label = "random normal values", 
                         value = "Histogram of Random Normal Values"),
		          plotOutput('hist'),
		          verbatimTextOutput('stats')
               )

server = function(input, output){
  data = reactive({
    rnorm(input$num)
  })
  output$hist = renderPlot({
    hist(data(), 
         main = isolate({input$title})) # isolate the refresh of the title text. 
  })
  output$stats = renderPrint({
    summary(data())
  })
}

shinyApp(ui = ui, server = server)
```

## Trigger event observer

start with an action button - actionButton(inputId = "clicks", label = "Click here")
uses observeEvent. 

```{r}
#observeEvent(input$clicks, {print(input$clicks)})
# ^function,  ^respond to change, ^run {this block of code} when observer is invalidated

ui = fluidPage(actionButton(inputId = 'clicks', 
                            label = "Touch Me"))

server = function (input, output){
  observeEvent(input$clicks, {print(as.numeric(input$clicks))})

}

shinyApp(ui = ui, server = server)
```
action buttons should just be used to start another action, the number of button clicks should not be used (like above)

more about action buttons: http://shiny.rstudio.com/articles/action-buttons/html

usually beyond what the user sees. Connect to database other things..
observeEvent(input$clicks, {code chunk})
here input$clicks is the reactive values to respond to, only when this value changes

### observe() 

observe() is similar syntax to render*, reactive, and isolate. Will run a block of code whenever a reactive value in the block of code changes (e.g., input$clicks is in there and updates, now the whole block will run)

## Delay reactions:
syntax: data = eventReactive(input$go, {rnorm(input$num)}) 
                              ^ respond to; ^ code use to (re)build object with invalidated input. 
                              
Example: want to add an Update button to app:
```{r}
ui = fluidPage(
               sliderInput(inputId = "num", 
                           label = "Choose a number", 
                           value = 25, min = 1, max = 100),
               actionButton(inputId = 'go', label = "Update"),
		          plotOutput('hist'),
		          verbatimTextOutput('stats')
               )

server = function(input, output){
  data = eventReactive(input$go,{
    rnorm(input$num)
  })
  output$hist = renderPlot({
    hist(data()) 
  })
  output$stats = renderPrint({
    summary(data())
  })
}

shinyApp(ui = ui, server = server)
```


## Manage State:
reactiveValues()

you can't change input values from the user. 
You can make your own reactive values that the user did not create. 

```{r}
ui = fluidPage(
  actionButton(inputId = 'norm', label = "Normal"),
  actionButton(inputId = 'unif', label = 'Uniform'),
  plotOutput("hist")
  )

server = function(input, output) {
  rv = reactiveValues(data = rnorm(100))
  
  observeEvent(input$norm, {rv$data = rnorm(100) })
  observeEvent(input$unif, {rv$data = runif(100) })
  
  output$hist = renderPlot({
    hist(rv$data)
  })
}

shinyApp(ui = ui, server = server )
```

here each button overwrites the rv$data function. 

Tips:
Keep the end result in mind. 
Code outside the server function will only be run once per R session(worker instance). So things like library() or loading a file that only needs to be loaded once should be outside the server function. 

Code inside the server will be run once per end user (connection). So code inside the server function will be run far more often, as needed by end user. 

Code inside a render function will be run once per reaction (many times). 

End Part 2 ##

# Part 3
How to style shiny app:
Look at shiny showcase. 
HTML is the code behind the UI, the shiny code just calls HTML code. 

Adding Static Content:
use tags() to change content. Tags is a list of functions

```{r}
names(tags)
# use () to run tag function
tags$h1()
# a tag is a link, first part is the link and second part is the hiyperlink text
tags$a(href= "www.rstudio.com", "RStudio")

# h1 first level header, h2, h3 and so on to h6. 
tags$h3("third level")
# plain text does not need a tag
fluidPage(
  "just plain text here"
)

tags$p("creates as a separate paragraph")

tags$em("italics")
tags$strong("Bold text")
tags$code("prints in a code block")
# nest tags together
fluidPage(
  tags$p("this is a", tags$strong("Shiny"), "app")
)
tags$br() # line break
tags$hr() # makes horizontal rule or line
# insert images
fluidPage(
  tags$img(height = 100, width = 100, 
           src = "http://www.rstudio.com/images/RStudio.2x.png")
) # src can be a url to an image, or include it your app


```
To add an image from a local file to the app, create and save files to subdir called "www". then you just need to src = 'filename.png'. only need subdir name if there is a subdir in the www folder

```{r}
# some tags have prebuild wrapper functions, so tags$ does not need to be called. Examples:
a() # hyper link
br() # line break
code() # code block
em() # italic
h1:h6() # header font sizes
hr() # horizontal line
img() # import image
p() # new paragraph
strong() # bold

#Example
fluidPage(
  h1("this damn app"), 
  p(style = 'font-family:Impact', "See other crap in the", 
    a(href = "http://www.rstudio.com/products/shiny/shiny-user-showcase/", "Shiny Shitty Showcase"))
)

```
if you want to pass raw HTML with HTML("HTML code here"), do this within fluidPage()

## Create a layout
place object based on x,y,z coornidates 
fluidRow() each time called will add a new row
works with column()

```{r}
fluidPage(
  fluidRow(
    column(3, offset = 8, sliderInput()), 
# offset and width are a base unit with max = 12 (left to right). The number just inside column(#) is the width. the width and offset add together and max at 12. 
  fluidRow(
    column(4, offset = 8,
           plotOutput("hist"))
  ))
)
```

## stopped at layers 1:59:00
